<style>
  .report-header {
    border-bottom: 2px solid #e5e5e5;
    padding-bottom: 16px;
    margin-bottom: 28px;
  }
  .report-title {
    font-size: 28px;
    font-weight: 700;
    margin-bottom: 6px;
  }
  .report-subtitle {
    font-size: 16px;
    color: #555;
    margin-bottom: 10px;
  }
  .report-meta {
    font-size: 14px;
    line-height: 1.6;
    color: #333;
  }
  .report-meta strong {
    font-weight: 600;
  }
  .repo-link a {
    color: #1a73e8;
    text-decoration: none;
  }
  .repo-link a:hover {
    text-decoration: underline;
  }
</style>

<div class="report-header">

<div class="report-title">
Social Perception of Body Morphology in Sports Students   
Pre-manipulation study
</div>

<div class="report-subtitle">
Master of Science (MSc) — Engineering and Ergonomics of Physical Activity  
Faculty of Sports and Health Sciences (UFR STAPS)  
University of Montpellier
</div>

<div class="report-meta">
<strong>Author:</strong> Delphine Hund (Master 1 IEAP Digimove)   
<strong>Supervisors:</strong> Ludovic Marin, Mathilde Parisi, Stéphane Raffard   
<strong>Course:</strong> Python, R and Git for Data Analysis  
<strong>Academic Year:</strong> 2025–2026  
<strong>Date of Defense:</strong> Tuesday 17 March 2026  
</div>

<div class="report-meta repo-link">
<strong>GitHub Repository:</strong>  
<a href="https://github.com/DelphineHun-student/Hund.Delphine-Project" target="_blank">
https://github.com/DelphineHun-student/Hund.Delphine-Project
</a>
</div>
<div style="
  margin-top:18px;
  font-size:16px;
  font-weight:600;
  color:#5a2d82;
  letter-spacing:0.5px;
">
Statistical Analysis Report
</div>

</div>

### What this document does

This `main.Rmd` is the **entry point** of the project. It runs the full pipeline:

1. **Preprocessing** (`PRG/01_preprocessing.R`)  
   → reads raw questionnaire data and exports a cleaned dataset into `DAT/clean/`

2. **Statistical report** (`PRG/02_statistics.Rmd`)  
   → renders an HTML report into `RES/`
   
3. **Plots** (`PRG/02_statistics.Rmd`)  
   → renders .png exports into `RES/plots_avatar/`

### How to run

- In RStudio: open the file and click on **Knit**

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  results = "hide"
)
```

### Restore the renv environment
```{r restore renv environment}
if (requireNamespace("renv", quietly = TRUE)) {
  renv::restore(prompt = FALSE)
}
```

### Run full pipeline
```{r run pipeline}
source("PRG/01_preprocessing.R")

rmarkdown::render(
  input = "PRG/02_statistics.Rmd",
  output_format = "html_document",
  output_dir = "RES")
```

```{r done, results='asis', echo=FALSE}
cat("### Outputs\n")
cat("- **Clean data:** `DAT/clean/`\n\n")
cat("- **Statistical report:** `RES/02_statistics.html`\n")
cat("- **Plots:** `RES/plots_avatar/`\n\n")
cat("✔ Full pipeline successfully executed.\n")
```