---
output:
  html_document:
    includes:
      before_body: assets/header.html
---

<div style="
  margin-top:18px;
  font-size:16px;
  font-weight:600;
  color:#5a2d82;
  letter-spacing:0.5px;
">
main.Rmd — Full pipeline execution
</div>

</div>

### What this document does

This `main.Rmd` is the **entry point** of the project. It runs the full pipeline:

1. **Preprocessing 1** (`PRG/01_preprocessing.R`)  
   → reads raw questionnaire data and exports a cleaned dataset into `DAT/clean/`

2. **Preprocessing 2** (`PRG/01.1_preprocessing.R`)  
   → runs the second preprocessing pipeline and exports its cleaned dataset (new) into `DAT/clean/`

3. **Statistical report 1** (`PRG/02_statistics.Rmd`)  
   → renders an HTML report into `RES/`

4. **Statistical report 2** (`PRG/02.1_statistics.Rmd`)  
   → renders a second HTML report into `RES/`
   
5. **Plots** (`PRG/02_statistics.Rmd`)  
   → renders .png exports into `RES/`

### How to run

- In RStudio: open the file and click on **Knit**

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  results = "hide"
)
```

```{r}
required_packages <- c("gt", "webshot2")

for (pkg in required_packages) {

  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg)
  }
}
```

### Restore the renv environment
```{r restore renv environment}
if (!requireNamespace("renv", quietly = TRUE)) {
  install.packages("renv")
}
source("renv/activate.R", local = TRUE)

renv::restore(prompt = FALSE)
```

### Run full pipeline
```{r run pipeline}
# Preprocessing 1
source("PRG/01_preprocessing.R")

# Preprocessing 2
source("PRG/01.1_preprocessing.R")

# Report 1
rmarkdown::render(
  input = "PRG/02_statistics.Rmd",
  output_format = "html_document",
  output_dir = "RES"
)

# Report 2
rmarkdown::render(
  input = "PRG/02.1_statistics.Rmd",
  output_format = "html_document",
  output_dir = "RES"
)
```
source("PRG/01_preprocessing.R")

source("PRG/01.1_preprocessing.R")

rmarkdown::render(
  input = "PRG/02_statistics.Rmd",
  output_format = "html_document",
  output_dir = "RES")

rmarkdown::render(
  input = "PRG/02.1_statistics.Rmd",
  output_format = "html_document",
  output_dir = "RES")


```{r done, results='asis', echo=FALSE}
cat("### Outputs\n")
cat("- **Clean data:** `DAT/clean/`\n")
cat("- **Statistical report:** `RES/02_statistics.html`\n")
cat("- **Plots (descriptive):** `RES/plots_avatar/`\n")
cat("- **Plots (ANOVA):** `RES/plots_avatar/`\n\n")
cat("✔ Full pipeline successfully executed.\n")
```