---
title: "Body perception pre-manip V2 : statistical analyses"
author: "Delphine Hund"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
    number_sections: true
    df_print: paged
---

<strong> Experiment summary </strong>

This pre-manipulation aims to validate the perceived morphology of dynamic avatars. Participants viewed 12 animated avatars (randomized order) and rated each avatar on 6 adjective-based dimensions: athletic, average, overweight, obese, thin, skinny.

The mapping between ratings and avatars is reconstructed using:
- VI03_01 ... VI03_12: the avatar identifier presented at each position (slot)
- PM01_01 ... PM72_01: 72 ratings corresponding to 12 blocks of 6 adjectives (fixed order within each block)

<strong> Hypotheses </strong>

For each avatar, ratings will be higher on the adjective matching its intended morphology than on the non-matching adjectives (congruent > incongruent). 

## Initiate
### Packages 
```{r}
easypackages::libraries("tidyverse", "DescTools", "readxl", "purrr","emmeans",
                        "dplyr", "tidyr", "stringr", "openxlsx", "ggplot2")
```

### Set directory 
```{r}
PRG_PATH = dirname(rstudioapi::getSourceEditorContext()$path) 
PATH = dirname(PRG_PATH)
DAT_PATH = file.path(PATH, "DAT")
CLEAN_PATH = file.path(DAT_PATH, "clean")
RES_PATH = file.path(PATH, "RES")
```

### Read data 
```{r}
file_rdata <- file.path(CLEAN_PATH, "pre_manip_ratings.rdata")
load(file_rdata)
```

## Data preparation
### Data integrity check
We verify that each participant contributes the expected number of ratings (12 avatars × 6 adjectives = 72).
```{r check rows}
ratings_long %>%
  count(CASE) %>%
  summarise(min_n = min(n), max_n = max(n))
```
### Variables types
We explicitly define variable types to ensure correct statistical modeling.
```{r variable type definition}
ratings_long <- ratings_long %>%
  mutate(
    CASE = as.factor(CASE), # Participant identifier
    avatar = as.character(avatar), # Avatar label
    adjectif = as.factor(adjectif), # Within-subject factor
    note = as.numeric(note), # Rating value
    DE03x01 = as.numeric(DE03x01), # Age
    DE08 = as.numeric(DE08) # Gender code (1 = Female ; 2 = Male ; 3 = Other)
  )
```

### Exclusion
Participants reporting significant eating disorder or body dysmorphic symptoms (DE10 = 2 or 3) were excluded from further analyses.
```{r exclusion participant ; DE10 (TCA/BDD)}
excluded_participants <- ratings_long %>%
  filter(DE10 %in% c(2, 3)) %>%
  distinct(CASE)

n_total <- ratings_long %>%
  distinct(CASE) %>%
  nrow()

n_excluded <- nrow(excluded_participants)
n_included <- n_total - n_excluded

ratings_clean <- ratings_long %>%
  filter(!(CASE %in% excluded_participants$CASE))

exclusion_summary <- tibble(
  Category = c("Total participants", 
               "Excluded (DE10 = 2 or 3)", 
               "Included in analyses"),
  N = c(n_total, n_excluded, n_included)
)
exclusion_summary
```

## Descriptive statistics
### Demographics: sample description
Age (mean ± SD) and gender distribution for the final included sample.
```{r demographic summary}
demographics <- ratings_clean %>%
  distinct(CASE, DE03x01, DE08)

demographics_summary <- demographics %>%    
  summarise(
    N_total = n(),
    Mean_age = round(mean(DE03x01, na.rm = TRUE), 2),
    SD_age = round(sd(DE03x01, na.rm = TRUE), 3),
    N_female = sum(DE08 == 1, na.rm = TRUE),
    N_male   = sum(DE08 == 2, na.rm = TRUE),
    N_other  = sum(DE08 == 3,na.rm = TRUE)
  )

demographics_summary
```

### Ratings by avatar × adjective
We report mean and standard deviation of ratings for each avatar across the six adjective dimensions.
```{r rating by avatar x adjective}
desc_avatar_adj <- ratings_clean %>%
  group_by(avatar, adjectif) %>%
  summarise(
    Mean = mean(note),
    SD = round(sd(note, na.rm = TRUE), 2),
    N = n(),
    .groups = "drop"
  ) %>%
  arrange(avatar, adjectif)

desc_avatar_adj
```

## Manip-check
### Congruence code
We extract each avatar’s intended morphology from its filename (e.g., `_ATH_`, `_AVG_`) and define whether each rating is *congruent* (adjective matches intended morphology) or *incongruent*.
```{r congruence code}
morpho_levels <- c("athletic","average","overweight","obese","thin","skinny")

ratings_clean <- ratings_clean %>%
  mutate(
    morpho_target = case_when(
      str_detect(avatar, "_ATH_")    ~ "athletic",
      str_detect(avatar, "_AVG_")    ~ "average",
      str_detect(avatar, "_OW_")     ~ "overweight",
      str_detect(avatar, "_OBESE_")  ~ "obese",
      str_detect(avatar, "_THIN_")   ~ "thin",
      str_detect(avatar, "_SKINNY_") ~ "skinny",
      TRUE ~ NA_character_
    ),
    morpho_target = factor(morpho_target, levels = morpho_levels),
    congruence = ifelse(adjectif == morpho_target, "congruent", "incongruent"),
    congruence = factor(congruence, levels = c("incongruent", "congruent"))
  )
# Check
table_morpho <- table(ratings_clean$morpho_target, useNA = "ifany")
table_congruence <- table(ratings_clean$congruence, useNA = "ifany")

table_morpho
table_congruence
```
### RM-ANOVA
For each avatar, we first test whether there is a global effect of adjective using a repeated-measures ANOVA (within-subject factor: adjective).  
```{r}
adjectifs <- levels(ratings_clean$adjectif)

# RM-ANOVA per avatar (global effect of adjective)
run_rm_anova <- function(data_av){ # data_av = data of ONE avatar
  data_av <- data_av %>% mutate(CASE = as.factor(CASE))

  aov_mod <- aov(note ~ adjectif + Error(CASE/adjectif), data = data_av)
  aov_sum <- summary(aov_mod)

  # Extract the adjective effect (within-subject)
  aov_table <- tryCatch({
    tab <- as.data.frame(aov_sum[["Error: CASE:adjectif"]][[1]]) %>%
      tibble::rownames_to_column("term") %>%
      filter(term == "adjectif") %>%
      transmute(
        df1 = Df,
        df2 = aov_sum[["Error: CASE:adjectif"]][[1]]["Residuals","Df"],
        F = `F value`,
        p_value = `Pr(>F)`
      )
    tab
  }, error = function(e){
    tibble(df1 = NA, df2 = NA, F = NA, p_value = NA)
  })
  
  aov_table
}
```

### Planned contrasts
We then conduct planned contrasts comparing the target adjective (i.e., the adjective corresponding to the avatar’s intended morphology) against each of the five non-target adjectives.

All post-hoc comparisons are corrected for multiple testing using the Holm method.
```{r}
run_planned_contrasts <- function(data_av, adjectives = adjectifs) {
  target <- unique(data_av$morpho_target)
  stopifnot(length(target) == 1)
  
  lm_mod <- lm(note ~ CASE + adjectif, data = data_av)
  emm <- emmeans(lm_mod, ~ adjectif)

  others <- adjectifs[adjectifs != target]

  contr_list <- setNames(
    lapply(others, function(adj){
      v <- rep(0, length(adjectifs)); names(v) <- adjectifs
      v[target] <- 1
      v[adj] <- -1
      v
    }),
    paste0(target, " - ", others)
  )

  ph <- contrast(emm, contr_list) %>%
    as.data.frame() %>%
    transmute(
      compared_to = str_replace(contrast, paste0("^", target, " - "), ""),
      estimate = estimate,
      SE = SE,
      df = df,
      t = t.ratio,
      p_value = p.value
    ) %>%
    mutate(
      p_adj = p.adjust(p_value, method = "holm"),
      significant = p_adj < .05
    ) 
  ph
}

out <- ratings_clean %>%
  group_by(avatar) %>%
  group_modify(~{
    data_av <- .x
    target <- unique(data_av$morpho_target)

    tibble(
      morpho_target = target,
      anova = list(run_rm_anova(data_av)),
      posthoc = list(run_planned_contrasts(data_av))
    )
  }) %>%
  ungroup()

anova_by_avatar <- out %>%
  tidyr::unnest(anova) %>%
  select(avatar, morpho_target, df1, df2, F, p_value)

posthoc_by_avatar <- out %>%
  tidyr::unnest(posthoc) %>%
  select(avatar, morpho_target, compared_to, estimate, t, p_value, p_adj, significant)

anova_by_avatar
posthoc_by_avatar
```

### Validation summary
```{r}
avatar_validation <- posthoc_by_avatar %>%
  group_by(avatar) %>%
  summarise(all_contrasts_significant = all(significant), .groups = "drop")

avatar_validation
```
## Plots and export
```{r}
library(dplyr)
library(ggplot2)
library(purrr)
library(stringr)

# 1) Moyennes + SEM par avatar × adjectif
plot_data <- ratings_clean %>%
  group_by(avatar, adjectif, morpho_target) %>%
  summarise(
    mean_note = mean(note, na.rm = TRUE),
    sem = sd(note, na.rm = TRUE) / sqrt(sum(!is.na(note))),
    .groups = "drop"
  )

# 2) Labels de significativité basés sur les post-hoc (Holm corrigé)
sig_labels <- posthoc_by_avatar %>%
  mutate(label = ifelse(significant, "*", "ns")) %>%
  left_join(
    plot_data,
    by = c("avatar", "morpho_target", "compared_to" = "adjectif")
  ) %>%
  mutate(
    y_pos = mean_note + sem + 2  # ajuste si les labels se chevauchent
  )

# 3) Fonction plot 1 avatar
plot_one_avatar <- function(av_id){

  pd_av <- plot_data %>% filter(avatar == av_id)
  sg_av <- sig_labels %>% filter(avatar == av_id)

  ggplot(pd_av, aes(x = adjectif, y = mean_note)) +
    geom_col(aes(fill = adjectif == morpho_target), width = 0.7) +
    geom_errorbar(aes(ymin = mean_note - sem, ymax = mean_note + sem), width = 0.2) +
    geom_text(
      data = sg_av,
      aes(x = compared_to, y = y_pos, label = label),
      size = 4
    ) +
    scale_fill_manual(
      values = c("grey70", "steelblue"),
      labels = c("Incongruent", "Congruent"),
      name = ""
    ) +
    labs(
      title = paste("Avatar:", av_id),
      subtitle = "RM-ANOVA + planned contrasts (congruent vs each incongruent, Holm-corrected)",
      x = "Adjective",
      y = "Mean rating"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "top"
    )
}

# 4) Afficher tous les avatars
avatar_list <- sort(unique(plot_data$avatar))
walk(avatar_list, ~ print(plot_one_avatar(.x)))

PLOTS_DIR <- file.path(RES_PATH, "plots_avatar")
if (!dir.exists(PLOTS_DIR)) dir.create(PLOTS_DIR, recursive = TRUE)

avatar_list <- sort(unique(plot_data$avatar))

# Export — un fichier PNG par avatar
walk(avatar_list, function(av){
  p <- plot_one_avatar(av)

  safe_name <- str_replace_all(av, "[^A-Za-z0-9_-]", "_")
  out_png <- file.path(PLOTS_DIR, paste0("plot_", safe_name, ".png"))

  ggsave(
    filename = out_png,
    plot = p,
    width = 10, height = 6, units = "in",
    dpi = 300, bg = "white"
  )
})

```

